---
title: "R Notebook"
output: html_notebook
---


### Example from the vignettes
```{r}
library(GERGM)
data("lending_2005")
data("covariate_data_2005")
data("net_exports_2005")

formula <- lending_2005 ~ edges + mutual(alpha = 0.8) + sender("log_GDP") + 
  receiver("log_GDP") + nodemix("G8", base = "No") + netcov(net_exports_2005) 

test <- GERGM::gergm(formula,
              covariate_data = covariate_data_2005,
              number_of_networks_to_simulate = 40000,
              thin = 1/100,
              proposal_variance = 0.05,
              MCMC_burnin = 10000,
              seed = 456,
              convergence_tolerance = 0.5)
```

### Reading the data and formatting it for the gergm package
```{r}

covariates <- read.table("data_cleaned/covariates.csv", sep = ";", row.names = 1, header = TRUE)
borders <- read.table("data_cleaned/borders.csv", sep = ",", row.names = 1, header = TRUE)
migration <- read.table("data_cleaned/migration.csv", sep = ",", row.names = 1, header = TRUE)

# Create borders adjacency matrix
mat<-matrix(0, 33, 33)
borders_am <- data.frame(mat, row.names = rownames(covariates))
colnames(borders_am) <- rownames(covariates)
for(i in 1:NROW(borders)) borders_am[ borders[i,1], borders[i,2] ] <- 1 

# Create migration adjacency matrix
mat<-matrix(0, 33, 33)
migration_am <- data.frame(mat, row.names = rownames(covariates))
colnames(migration_am) <- rownames(covariates)
for(i in 1:NROW(migration)) migration_am[ migration[i,1], migration[i,2] ] <- migration[i,3]

# Manually fixing up the data a bit more...
# Deleting Korea and Slovak Republic because they have NAs
migration_am <- migration_am[-c(34, 35), -c(34, 35)]
# Also deleting Chile, Greece, Ireland and Portugal because too many missing values
migration_am <- migration_am[-c(5, 12, 15, 26), -c(5, 12, 15, 26)]
# Adding a bunch of borders that are missing from the dataset
# TODO
# Changing zeroes to like 1 or 0.05 if it is good for the model?
# TODO

# Converting the adjacency matrix into matrix type
migration_mat <- as.matrix(migration_am)
# Making sure the diagonals stay zero (for when we are raising values)
diag(migration_mat) <- 0

GERGM::plot_network(migration_mat, white_background = TRUE, show_legend = FALSE)
```


### Scaling the covariates to similar ranges for more readable coefficients
```{r}
covariates <- subset(covariates, select = -c(Area, Safety_Index))
summary(covariates)

# Population is now on the scale of 1 million inhabitants
covariates$Population <- covariates$Population / 1000000
# GDP will be a 100th of the GDP...
covariates$GDP <- covariates$GDP / 100
# Crime index will get a zero at the end
covariates$Crime_Index <- covariates$Crime_Index * 10

summary(covariates)
```


### Loading data into a first gergm model
```{r}
formula1 <- migration_mat ~ edges + sender("Population")

model1 <- GERGM::gergm(formula1,
                       covariate_data = covariates,
                   number_of_networks_to_simulate = 40000,
                   thin = 1/100,
                   proposal_variance = 0.05,
                   MCMC_burnin = 10000,
                   seed = 456,
                   convergence_tolerance = 0.5)

```


### Second model with "mutual"
```{r}
formula2 <- migration_mat ~ edges + mutual

model2 <- GERGM::gergm(formula2,
                       covariate_data = covariates,
                       number_of_networks_to_simulate = 40000,
                       thin = 1/100,
                       proposal_variance = 0.05,
                       MCMC_burnin = 10000,
                       seed = 456,
                       convergence_tolerance = 0.5,
                       maximum_number_of_theta_updates = 10)
# verbose = FALSE
# maximum_number_of_theta_updates = 20
```


### Third model with a few cherry-picked different effects
```{r}

formula3 <- migration_mat ~ 
                            edges + 
                            sender("Population") + 
                            sender("Crime_Index") + 
                            in2stars(alpha = 1) + 
                            nodematch("Language")

model3 <- GERGM::gergm(formula3,
                       covariate_data = covariates,
                       number_of_networks_to_simulate = 40000,
                       thin = 1/100,
                       proposal_variance = 0.05,
                       MCMC_burnin = 10000,
                       seed = 456,
                       convergence_tolerance = 0.5,
                       maximum_number_of_theta_updates = 10)
# verbose = FALSE
# maximum_number_of_theta_updates = 20
```


### Fourth model with all exogenous effects
```{r}

formula4 <- migration_mat ~ 
                            edges + 
                            sender("Population") + 
                            receiver("Population") +
                            sender("Population_Density") +
                            receiver("Population_Density") +
                            sender("GDP") +
                            receiver("GDP") +
                            sender("Crime_Index") +
                            receiver("Crime_Index") +
                            nodematch("Language")

model4 <- GERGM::gergm(formula4,
                       covariate_data = covariates,
                       number_of_networks_to_simulate = 40000,
                       thin = 1/100,
                       proposal_variance = 0.05,
                       MCMC_burnin = 10000,
                       seed = 456,
                       convergence_tolerance = 0.5,
                       maximum_number_of_theta_updates = 10)
# verbose = FALSE
# maximum_number_of_theta_updates = 20
```


### Fifth model with all endogenous and exogenous effects
```{r}

formula5 <- migration_mat ~ 
                            edges + 
                            sender("Population") + 
                            receiver("Population") +
                            sender("Population_Density") +
                            receiver("Population_Density") +
                            sender("GDP") +
                            receiver("GDP") +
                            sender("Crime_Index") +
                            receiver("Crime_Index") +
                            nodematch("Language") +
                            in2stars(alpha = 0.8) +
                            out2stars(alpha = 0.8) +
                            mutual(alpha = 0.8)

model5 <- GERGM::gergm(formula5,
                       covariate_data = covariates,
                       number_of_networks_to_simulate = 40000,
                       thin = 1/100,
                       proposal_variance = 0.05,
                       MCMC_burnin = 10000,
                       seed = 456,
                       convergence_tolerance = 0.5,
                       maximum_number_of_theta_updates = 10)
# verbose = FALSE
# maximum_number_of_theta_updates = 20
```

