---
title: "R Notebook"
output: html_notebook
---

### Importing data
```{r}
covariates <- read.table("data_final/covariates.csv", row.names = 1, sep = ",", header = TRUE)
borders <- as.matrix(read.table("data_final/borders_adjacency_matrix.csv",sep = ";", row.names = 1, header = TRUE))
migration <- as.matrix(read.table("data_final/migration_adjacency_matrix.csv", sep = ";", row.names = 1, header = TRUE))


```

### Plotting data
```{r}
GERGM::plot_network(migration, white_background = TRUE, show_legend = FALSE)
GERGM::plot_network(borders, white_background = TRUE, show_legend = FALSE)
```

### Scaling the data
```{r}
summary(covariates)
                                                                     #covariates <- subset(covariates, select = -c(Area, Safety_Index))

# You could do something like this to put covariates on a different scale like million inhabitants
# covariates$Population <- covariates$Population / 1000000
# Or this to try log scales like Claudia suggested
# covariates$Population <- log(covariates$Population)
```

### Cauchy model: 00
```{r}
formula <- migration ~ 
                        edges + 
                        sender("Population") + 
                        receiver("Population") +
                        sender("Population_Density") +
                        receiver("Population_Density") +
                        sender("GDP") +
                        receiver("GDP") +
                        sender("Crime_Index") +
                        receiver("Crime_Index") +
                        sender("Freedom_2020") +
                        receiver("Freedom_2020") +
                        sender("Unemployment_2020") +
                        receiver("Unemployment_2020") +
                        nodematch("Language") +
                        netcov(borders) +
                        in2stars(alpha = 0.8) +
                        out2stars(alpha = 0.8) +
                        mutual(alpha = 0.8)


model <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```
### Cauchy model lower convergence tolerance: 01
```{r}
model01 <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.3,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```

### Cauchy model higher convergence tolerance: 02
```{r}
model02 <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.7,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```


### Cauchy model lower proposal variance: 03
```{r}
model03 <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.03,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```

### DIDN'T WORK Cauchy model higher proposal variance (default): 04
```{r}
model04 <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.1,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```




### Exclude language 
```{r}
formula_no_language <- migration ~ 
                        edges + 
                        sender("Population") + 
                        receiver("Population") +
                        sender("Population_Density") +
                        receiver("Population_Density") +
                        sender("GDP") +
                        receiver("GDP") +
                        sender("Crime_Index") +
                        receiver("Crime_Index") +
                        sender("Freedom_2020") +
                        receiver("Freedom_2020") +
                        sender("Unemployment_2020") +
                        receiver("Unemployment_2020") +
                        netcov(borders) +
                        in2stars(alpha = 0.8) +
                        out2stars(alpha = 0.8) +
                        mutual(alpha = 0.8)


model_no_language <- GERGM::gergm(formula_no_language,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
```

### Cauchy model lower convergence tolerance: 11
```{r}
model11 <- GERGM::gergm(formula_no_language,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.3,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```

### Cauchy model higher convergence tolerance: 12
```{r}
model12 <- GERGM::gergm(formula_no_language,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.7,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```


### no_language model lower proposal variance: 13
```{r}
model13 <- GERGM::gergm(formula_no_language,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.03,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```

### DIDN'T WORK no_language model higher proposal variance (default): 14
```{r}
model14 <- GERGM::gergm(formula_no_language,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.1,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```

# Model 2: 100 000 networks (more)
```{r}
model2 <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 100000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```
# Model 3: 400 000 networks (more)
```{r}
model3 <- GERGM::gergm(formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 400000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      # hyperparameter_optimization = TRUE,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
beepr::beep()
```

Covariates & Summary
```{r}
data.frame(colnames(covariates))
# summary(attributes(model))
# # this one's OK
# gof <- GERGM::GOF(model, return_GERGM_Object = TRUE)
# gof01 <- GERGM::GOF(model01, return_GERGM_Object = TRUE)
# worse
# gof02 <- GERGM::GOF(model02, return_GERGM_Object = TRUE)
# gof03 <- GERGM::GOF(model03, return_GERGM_Object = TRUE)
# broken
# gof04 <- GERGM::GOF(model04, return_GERGM_Object = TRUE)
# gof10 <- GERGM::GOF(model_no_language, return_GERGM_Object = TRUE)
# the 6th graph is nicer
# gof11 <- GERGM::GOF(model11, return_GERGM_Object = TRUE)
# gof12 <- GERGM::GOF(model12, return_GERGM_Object = TRUE)
# gof13 <- GERGM::GOF(model13, return_GERGM_Object = TRUE)
# broken
# gof14 <- GERGM::GOF(model14, return_GERGM_Object = TRUE)
# HORRIBLE
# gof2 <- GERGM::GOF(model2, return_GERGM_Object = TRUE)
gof3 <- GERGM::GOF(model3, return_GERGM_Object = TRUE)
# beepr::beep()
```


# TODO after getting the best of the previous models, make a model with more networks simulated


SIGNIFICANCE BASED ON LOWER & UPPER SE 
```{r}
(EstSE <- rbind(t(attributes(model)$theta.coef), t(attributes(model)$lambda.coef)))
EstSE_df = as.data.frame(EstSE)

EstSE_df$LOWER = EstSE_df$est - EstSE_df$se*(-qnorm((1-0.95)/2))
EstSE_df$UPPER = EstSE_df$est + EstSE_df$se*(-qnorm((1-0.95)/2))

# If LOWER and UPPER have the same sign (positive/negative) then there's significance
EstSE_df$Significant = (EstSE_df$LOWER > 0 & EstSE_df$UPPER > 0) | (EstSE_df$LOWER < 0 & EstSE_df$UPPER < 0)

```
Non-zero model
```{r}
zeroless_migration <- migration
zeroless_migration[zeroless_migration==0] <- 0.01
diag(zeroless_migration) <- 0

zeroless_formula <- zeroless_migration ~ 
                        edges + 
                        sender("Population") + 
                        receiver("Population") +
                        sender("Population_Density") +
                        receiver("Population_Density") +
                        sender("GDP") +
                        receiver("GDP") +
                        sender("Crime_Index") +
                        receiver("Crime_Index") +
                        sender("Freedom_2020") +
                        receiver("Freedom_2020") +
                        sender("Unemployment_2020") +
                        receiver("Unemployment_2020") +
                        nodematch("Language") +
                        netcov(borders) +
                        in2stars(alpha = 0.8) +
                        out2stars(alpha = 0.8) +
                        mutual(alpha = 0.8)

```


GERGM LogCauchy transformation
```{r}
# (Cauchy is the default transformation, so we it is above)
model <- GERGM::gergm(zeroless_formula,
                      covariate_data = covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 7000, # try less
                      seed = 456,
                      convergence_tolerance = 0.5,
                      transformation_type = "LogCauchy",
                      maximum_number_of_theta_updates = 5, # try less
                      maximum_number_of_lambda_updates = 10)
```


Z-scores 
```{r}
# qnorm(p, mean = 0, sd = 1, lower.tail = TRUE, log.p = FALSE)
# p	vector of probabilities.
# qnorm gives the quantile function
# Can manually encode the p-values given in the previous cell
# OR should the z-score formula of each attribute be taken? 
# z_scores <- (data-mean(data))/sd(data)
zscore_Population = (covariates$Population - mean(covariates$Population))/sd(covariates$Population)

zscore_Area = (covariates$Area - mean(covariates$Area))/sd(covariates$Area)

zscore_Population_Density = (covariates$Population_Density - mean(covariates$Population_Density))/sd(covariates$Population_Density)

zscore_GDP = (covariates$GDP - mean(covariates$GDP))/sd(covariates$GDP)

zscore_Crime_Index = (covariates$Crime_Index - mean(covariates$Crime_Index))/sd(covariates$Crime_Index)

zscore_Safety_Index = (covariates$Safety_Index - mean(covariates$Safety_Index))/sd(covariates$Safety_Index)

zscore_Freedom_2020 = (covariates$Freedom_2020 - mean(covariates$Freedom_2020))/sd(covariates$Freedom_2020)

zscore_Unemployment_2020 = data.frame(covariates$Unemployment_2020 - mean(covariates$Unemployment_2020))/sd(covariates$Unemployment_2020)


z_covariates <- data.frame(zscore_Population, zscore_Area, zscore_Population_Density, zscore_GDP, zscore_Crime_Index, zscore_Safety_Index, zscore_Freedom_2020, zscore_Unemployment_2020)

colnames(z_covariates) <- colnames(covariates)[-7] # W/o freedom
rownames(z_covariates) <- rownames(covariates)
# This doesn't converge if we maintain the original GERGM inputs
```

```{r}
z_formula <- migration ~ 
                        edges + 
                        sender("Population") + 
                        receiver("Population") +
                        sender("Population_Density") +
                        receiver("Population_Density") +
                        sender("GDP") +
                        receiver("GDP") +
                        sender("Crime_Index") +
                        receiver("Crime_Index") +
                        sender("Freedom_2020") +
                        receiver("Freedom_2020") +
                        sender("Unemployment_2020") +
                        receiver("Unemployment_2020") +
                        netcov(borders) +
                        in2stars(alpha = 0.8) +
                        out2stars(alpha = 0.8) +
                        mutual(alpha = 0.8)


z_model <- GERGM::gergm(z_formula,
                      covariate_data = z_covariates,
                      number_of_networks_to_simulate = 40000,
                      thin = 1/100,
                      proposal_variance = 0.05,
                      MCMC_burnin = 10000,
                      seed = 456,
                      convergence_tolerance = 0.5,
                      maximum_number_of_theta_updates = 10,
                      maximum_number_of_lambda_updates = 10)
```


### Checking other fun stuff like details of the model
```{r}
# GERGM::Estimate_Plot(model,
#               coefficients_to_plot = "both",
#               leave_out_coefficients = "intercept")
# 
# xtable::xtable(EstSE, dig=4)
# 
# summary(attributes(model))
```









